\documentclass{article}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}

\begin{document}

\title{Programming for Performance: Assignment 1}
\author{Your Names}
\renewcommand{\today}{}
\maketitle

\section*{Application characteristics}

``characteristics of the software: language and platform, size of the codebase, how actively-maintained is the codebase, any design information you can find (does it follow a design pattern?), userbase''

We chose to work with FOO, which is available from \url{http://foo}.
FOO is an application for computing BAR. It is written in Scheme and
has 20,000 lines of code. (Write anything else you choose about
the application characteristics. You might say something about its
age, developer community, etc.)

\section*{Workload}

``a sample, reproducible workload which tests some aspect of the performance (you might want to put this on the Internet)''

``baseline performance data on this workload''

\section*{Profiling}
We ran the XYZ profiling tool on our application. Our workload was XXX. 
We found YYY.

``profiling results for the workload, and discussion of these results.''

\section*{Code Improvement}
Provide details of the system performance before and after your
change, including enough information about the workload that we can
reproduce your observations. I suggest that you hand in something like:
\begin{itemize}
\item diffs for your changes;
\item a discussion of why these changes improve performance; and
\item evidence that your changes do improve performance.
\end{itemize}

\end{document}
